\section{Fitting Active Appearance Models}
\label{sec:fit}
AAM fitting is typically formulated as a search over the shape and appearance parameters that minimize the \emph{Sum of Squared pixel Differences} (SSD) between the warped input image and the linear appearance model
\begin{align}
\mathbf{p}_o, \mathbf{c}_o & = \underset{\mathbf{p}, \mathbf{c}}{\mathrm{arg\,min\;}} 
\left\| \mathbf{i}[\mathbf{p}] - \left( \bar{\mathbf{a}} + \mathbf{A} \mathbf{c} \right) \right\|^2 
\label{eq:ssd}
\end{align}

Several techniques have been proposed to solve the non-linear minimization problem defined by Eq. \ref{eq:ssd}. In this paper, we will center the discussion around compositional gradient descent algorithms \cite{Matthews2004, Gross2005, Papandreou2008, Amberg2009, Tzimiropoulos2012, Tzimiropoulos2013} for fitting AAMs. Consequently, we will not review discriminative and regression based approaches. For these methods, we refer the interested reader to the existent literature \cite{Cootes2001, Hou2001, Batur2005, Donner2006, Liu2009, Saragih2009, Sauer2011, Xiong2013}.

\subsection{Compositional Gauss-Newton}
\label{sec:cgd}
Assuming, for the time being, that the true appearance parameters $\mathbf{c}_o$ are known, the problem defined by Eq. \ref{eq:ssd} reduces to a non-rigid image alignment problem \cite{Baker2004, Munoz2014} between a particular instance of the object present in the image and its optimal appearance reconstruction by the appearance model:
\begin{equation}
    \begin{aligned}
        \mathbf{p}_o & = \underset{\mathbf{p}}{\mathrm{arg\,min\;}} 
        \left\| \mathbf{i}[\mathbf{p}] - \mathbf{a} \right\|^2 
    \label{eq:ssd_shape}
    \end{aligned}
\end{equation}
where $\mathbf{a} = \bar{\mathbf{a}} + \mathbf{A} \mathbf{c}_o$ is obtained by directly evaluating Eq. \ref{eq:app_model} given the true appearance parameters $\mathbf{c}_o$.

Compositional Gauss-Newton algorithms solve the previous non-linear optimization problem with respect to the shape parameters $\mathbf{p}$ by:
\begin{enumerate}
    \item Introducing an incremental warp $\mathcal{W}(\mathbf{x}; \delta \mathbf{p})$; either on the image side:
    \label{it:step_1}
    \begin{equation}
        \begin{aligned}
            \delta \mathbf{p}_o & = \underset{\delta \mathbf{p}}{\mathrm{arg\,min\;}} \left\| \mathbf{i} [\mathbf{p} \circ \delta \mathbf{p}] - \mathbf{a} \right\|^2 
        \label{eq:fc}
        \end{aligned}
    \end{equation}
    in what is known as \emph{Forward Composition} (FC) or on the model side: 
    \begin{equation}
        \begin{aligned}
            \delta \mathbf{p}_o = \underset{\delta \mathbf{p}}{\mathrm{arg\,min\;}}  \left\| \mathbf{i} [\mathbf{p}] - \mathbf{a}[\delta \mathbf{p}] \right\|^2 
        \label{eq:ic}
        \end{aligned}
    \end{equation}
    in what is known as \emph{Inverse Composition} (IC).
    
    \item Performing a Taylor expansion of the residual around the incremental warp
    \begin{equation}
        \begin{aligned}
            \delta \mathbf{p}_o & = \underset{\delta\mathbf{p}}{\mathrm{arg\,min\;}} \left\| \mathbf{r}(\mathbf{p}) + \frac{\partial\mathbf{r}}{\partial\mathbf{p}} \delta\mathbf{p} \right\|^2
            \\
            & = \underset{\delta\mathbf{p}}{\mathrm{arg\,min\;}} \left\| \mathbf{i} [\mathbf{p}] - \mathbf{a} + \mathbf{J} \delta \mathbf{p} \right\|^2
        \label{eq:taylor}
        \end{aligned}
    \end{equation}
    where $\mathbf{J} = \nabla \mathbf{i}[\mathbf{p}] \frac{\partial\mathcal{W}}{\partial\mathbf{p}}$ or $\mathbf{J} = -\nabla \mathbf{a} \frac{\partial\mathcal{W}}{\partial\mathbf{p}}$ depending on whether forward or inverse composition is used.
    
    \item Solving for $\delta\mathbf{p}$ by equating the derivative of Eq. \ref{eq:taylor} to $\mathbf{0}$:
    \begin{equation}
        \begin{aligned}
            \frac{\partial \left\| \mathbf{r}(\mathbf{p}) + \frac{\partial\mathbf{r}}{\partial\mathbf{p}} \delta\mathbf{p} \right\|^2}{\partial \mathbf{p}} & = \mathbf{0} 
            \\
            \left(\mathbf{r}(\mathbf{p}) + \frac{\partial\mathbf{r}}{\partial\mathbf{p}} \delta\mathbf{p} \right) \frac{\partial\mathbf{r}}{\partial\mathbf{p}}^T  & = \mathbf{0} 
            \\
            \frac{\partial\mathbf{r}}{\partial\mathbf{p}}^T \mathbf{r}(\mathbf{p}) + \frac{\partial\mathbf{r}}{\partial\mathbf{p}}^T \frac{\partial\mathbf{r}}{\partial\mathbf{p}} \delta\mathbf{p} & = \mathbf{0} 
        \label{eq:taylor}
        \end{aligned}
    \end{equation}
    Rearranging, the optimal solution $\delta\mathbf{p}_o$ is given by:
    \begin{equation}
        \begin{aligned}
            \delta\mathbf{p}_o & = - \left( \frac{\partial\mathbf{r}}{\partial\mathbf{p}}^T \frac{\partial\mathbf{r}}{\partial\mathbf{p}} \right)^{-1} \frac{\partial\mathbf{r}}{\partial\mathbf{p}}^T \mathbf{r}(\mathbf{p})
            \\
            & = - \mathbf{H}^{-1} \mathbf{J}^T \left( \mathbf{i}[\mathbf{p}] - \mathbf{a} \right)
        \label{eq:solve_dp}
        \end{aligned}
    \end{equation}
    
    \item Updating the current estimate of the warp using one of the following update rules
    \begin{equation}
        \mathcal{W}(\mathbf{x}; \mathbf{p}_o) \leftarrow \mathcal{W}(\mathbf{x}; \mathbf{p}) \circ \mathcal{W}(\mathbf{x}; \delta \mathbf{p}_o) 
        \label{eq:fc_update}
    \end{equation}
    for forward composition and
    \begin{equation}
        \mathcal{W}(\mathbf{x}; \mathbf{p}_o) \leftarrow \mathcal{W}(\mathbf{x}; \mathbf{p}) \circ \mathcal{W}(\mathbf{x}; \delta \mathbf{p}_o)^{-1} 
        \label{eq:ic_update}
    \end{equation}
    for inverse.
    
    \item Going back to Step \ref{it:step_1} until: 
    \begin{equation}
        \begin{aligned}
            \left\| \mathbf{p}_{k+1} - \mathbf{p}_k \right\|^2 < \epsilon
        \end{aligned}
    \end{equation}
    where $\epsilon$ is a small number determining the convergence of the shape parameters $\mathbf{p}$.
\end{enumerate}
Further details on the computation of $\frac{\partial \mathcal{W}}{\partial \mathbf{p}}$, composition $\mathcal{W}(\mathbf{x}; \mathbf{p}) \circ \mathcal{W}(\mathbf{x}; \delta \mathbf{p})$ and inversion $\mathcal{W}(\mathbf{x};\delta \mathbf{p})^{-1}$ of typical AAM motion models such as PWA and TPS warps can be found in \cite{Matthews2004, Papandreou2008}.

In general, optimization of Eq. \ref{eq:ssd} with respect to the appearance parameters is highly dependent on the particular algorithm being used. The next subsections provide a detailed review of the main existent strategies used by compositional Gauss-Newton algorithms to deal with the appearance parameters.

\begin{table*}
\begin{tabular}{lccccccc}
\hline\noalign{\smallskip}
Algorithm& $\mathbf{a}$ & $\mathbf{J}$ & $\mathbf{J}^T\mathbf{Q}$ & $\mathbf{H}$ & $\mathbf{H}^{-1}$ & $\mathbf{J}^T\mathbf{e}$ & $\mathbf{H}^{-1}\mathbf{J}^T\mathbf{e}$ \\
\noalign{\smallskip}\hline\noalign{\smallskip}
PFC & - & $O(nF)$ & $O(mnF)$ & $O(n^2F)$ & $O(n^3)$ & $O(nF)$ & $O(n^2)$ \\
PIC & - & - & - & - & - & $O(nF)$ & - \\
SFC & $O(mF)$ & $O(nF)$ & - & $O((m+n)^2F)$ & $O((m+n)^3)$ & $O((m+n)F)$ & $O((m+n)^2)$ \\
SIC & $O(mF)$ & $O(nF)$ & - & $O((m+n)^2F)$ & $O((m+n)^3)$ & $O((m+n)F)$ & $O((m+n)^2)$ \\
AFC & $O(mF)$ & $O(nF)$ & - & $O(n^2F)$ & $O(n^3)$ & $O(nF)$ & $O(n^2)$ \\
AIC & $O(mF)$ & $O(nF)$ & - & $O(n^2F)$ & $O(n^3)$ & $O(nF)$ & $O(n^2)$ \\
Fast-SFC & - & $O(nF)$ & $O(mnF)$ & $O(n^2F)$ & $O(n^3)$ & $O(nF)$ & $O(n^2)$ \\
Fast-SIC &  $O(mF)$ & $O(nF)$ & $O(mnF)$ & $O(n^2F)$ & $O(n^3)$ & $O(nF)$ & $O(n^2)$ \\
\noalign{\smallskip}\hline
\end{tabular}
\caption{Computational cost per iteration of Compositional Gradient Descent algorithms for fitting AAMs.}
\label{tab:complexity}  
\end{table*}
